\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-11 tex/paper.tex
reworded problem statement

We assume that the stochastic generative process that defines $p(\mathbf{x}|\bfz,\bftheta)$ is
specified through a deterministic function $g(\cdot; \bftheta) : {\cal Z} \to
\mathbb{R}^d$ with usually few parameters. That is, we consider
\begin{equation}\label{eqn:p_theta}
    \mathbf{x} \sim p(\mathbf{x}|\bftheta) \equiv \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta)
\end{equation}
such that the likelihood $p(\mathbf{x}|\bftheta)$ can be 
considered as a non-differentiable reparameterization of the random variable $\bfz$
\begin{equation}\label{eqn:p_x_sim}
    p(\mathbf{x}|\bftheta) = \frac{\partial}{\partial x_1} \dots \frac{\partial}{\partial x_d} \int_{\{\bfz:g(\bfz;\bftheta) \leq \mathbf{x}\}} p(\bfz|\bftheta) \mu(d\bfz),
\end{equation}
where $\mu$ is a probability measure.
%Importantly, the simulator $g$ is assumed to be a non-invertible function, that can only be
%used to generate data in forward mode. For this reason, evaluating the integral
%in Eqn.~\ref{eqn:p_x_sim} is intractable. As commonly found
%in science, we finally assume the lack of access to or existence of derivatives of $g$ with respect to $\bftheta$,
%e.g. as when $g$ is specified as a computer program.

Given some observed data $\{ \mathbf{x}_i | i=1, \dots, N \}$ drawn from the
(unknown) true distribution $p_r(\mathbf{x})$, our goal is the inference of the parameters
of interest $\bftheta^*$ that minimize the divergence between $p_r(\mathbf{x})$ and
the modeled data distribution $p(\mathbf{x}|\bftheta)$ induced by $g(\cdot;
\bftheta)$ over $\bfz$. That is,
\begin{equation}
    \bftheta^* = \arg \min_{\bftheta \in \bftheta} \rho(p_r(\mathbf{x}), p(\mathbf{x}|\bftheta)),
\end{equation}
where $\rho$ is some distance or divergence.

>>>>

We assume that the stochastic generative process that defines $p(\mathbf{x}|\bftheta)$ is
specified through a non-differentiable deterministic function $g(\cdot; \bftheta) : {\cal Z} \to
\mathbb{R}^d$. Operationally, %That is, we consider
\begin{equation}\label{eqn:p_theta}
    \mathbf{x} \sim p(\mathbf{x}|\bftheta) \equiv \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta)
\end{equation}
such that the density $p(\mathbf{x}|\bftheta)$ can be 
rewritten as
%considered as a non-differentiable reparameterization of the random variable $\bfz$
%\begin{equation}\label{eqn:p_x_sim}
%    p(\mathbf{x}|\bftheta) = \frac{\partial}{\partial x_1} \dots \frac{\partial}{\partial x_d} \int_{\{\bfz:g_i(\bfz;\bftheta) \leq x_i\}} p(\bfz|\bftheta) \mu(d\bfz),
%\end{equation}
\begin{equation}\label{eqn:p_x_sim}
    p(\mathbf{x}|\bftheta) = \int_{\{\bfz:g(\bfz;\bftheta) = \bfx \}} p(\bfz|\bftheta) \mu(d\bfz),
\end{equation}
where $\mu$ is a probability measure.
%Importantly, the simulator $g$ is assumed to be a non-invertible function, that can only be
%used to generate data in forward mode. For this reason, evaluating the integral
%in Eqn.~\ref{eqn:p_x_sim} is intractable. As commonly found
%in science, we finally assume the lack of access to or existence of derivatives of $g$ with respect to $\bftheta$,
%e.g. as when $g$ is specified as a computer program.

Given some observed data $\{ \mathbf{x}_i | i=1, \dots, N \}$ drawn from the
(unknown) true distribution $p_r(\mathbf{x})$, our goal is to estimate the parameters 
$\bftheta^*$ that minimize the divergence between $p_r(\mathbf{x})$ and
the implicit model $p(\mathbf{x}|\bftheta)$. That is,
\begin{equation}
    \bftheta^* = \arg \min_{\bftheta \in \bftheta} \rho(p_r(\mathbf{x}), p(\mathbf{x}|\bftheta)),
\end{equation}
where $\rho$ is some distance or divergence.
